### 2013-05-17 draft.tex
fixed the clumped isotope analysis section of the methods, still may need to be changed to reflect calculation of ETF

Each sample was measured between one and three times. Approximately 10 milligrams of powdered sample were dissolved in phosphoric acid at 90\degrees C, which produces CO$_2$ with a known oxygen isotopic offset from the sample carbonate. The extraction and subsequent cleaning of the CO$_2$ followed procedures outlined by \cite{Ghosh2006}. The cleaned CO$_2$ was analyzed on a Finnigan MAT253 gas source mass spectrometer configured to collect masses 44 through 49. The sample gas was measured in eight separate acquisitions, each of which was bracketed by a measurement of reference CO$_2$ with known \deltac C and \deltao O. The \deltac C, \deltao O, $\delta_{45}$, $\delta_{46}$, $\delta_{47}$, and $\delta_{48}$ were calculated for each sample using the reference CO$_2$ as the reference material. Acquisition-to-acquisition standard deviations were on average 0.03$\permil$ for $\delta_{47}$, 0.04$\permil$ for \deltao O, and 0.02$\permil$ for \deltac C. Preliminary values for $\Delta_{47}$ and $\Delta_{48}$ of each sample were also calculated using the reference CO$_2$ as the reference material and the following equations:

\begin{align}
\Delta_{47\text{ prelim}}&=\left[\left(\frac{\text{R}^{47}_{\text{sa}}}{\text{R}^{47*}_{\text{sa}}}\frac{\text{R}^{47*}_{\text{ref}}}{\text{R}^{47}_{\text{ref}}}-1\right)-\left(\frac{\text{R}^{46}_{\text{sa}}}{\text{R}^{46*}_{\text{sa}}}\frac{\text{R}^{46*}_{\text{ref}}}{\text{R}^{46}_{\text{ref}}}-1\right)-\left(\frac{\text{R}^{45}_{\text{sa}}}{\text{R}^{45*}_{\text{sa}}}\frac{\text{R}^{45*}_{\text{ref}}}{\text{R}^{45}_{\text{ref}}}-1\right)\right]\times 1000\\
\Delta_{48\text{ prelim}}&=\left[\left(\frac{\text{R}^{48}_{\text{sa}}}{\text{R}^{48*}_{\text{sa}}}\frac{\text{R}^{48*}_{\text{ref}}}{\text{R}^{48}_{\text{ref}}}-1\right)-2\left(\frac{\text{R}^{46}_{\text{sa}}}{\text{R}^{46*}_{\text{sa}}}\frac{\text{R}^{46*}_{\text{ref}}}{\text{R}^{46}_{\text{ref}}}-1\right)\right]\times 1000
\end{align}

\noindent where R* refers to the stochastic distribution of masses calculated from the bulk isotopic composition of the sample (sa) or reference gas (ref).

Measurements were standardized by comparison to heated gases, which are CO$_2$ with two different bulk isotopic compositions that was heated at 1000\degrees C for three hours, and then cooled rapidly to room temperature. At least one heated gas was measured each day, and each sample was compared only to the heated gases that were run during the seven days before and after the date on which the sample was measured. The $\delta_{48}$ and $\Delta_{48\text{ prelim}}$ of each sample was compared to a line of best fit through a plot of $\delta_{48}$ versus $\Delta_{48}$ of the heated gases. All samples whose $\Delta_{48}$ fell more than 1$\permil$ off the line were discarded due to likely contamination. The $\Delta_{47}$ values of the remaining samples were corrected so that the $\Delta_{47}$ reflects a comparison to the stochastic heated gases, rather than the reference gas, using the following equation:

\begin{align}
\Delta_{47}=\Delta_{47\text{ prelim}}-\left(\delta^{47}\cdot\text{M}_{\text{HG}}+\text{b}_{\text{HG}}\right)\cdot\text{K}_{\text{stretch}}+\text{K}_{\text{acid}}
\end{align}

\noindent where M$_{\text{HG}}$ is the slope of the best fit line through a plot of $\delta_{47}$ versus $\Delta_{47}$ of heated gases, b$_{\text{HG}}$ is the intercept of the same line, K$_{\text{stretch}}=-0.8453$ and K$_{\text{acid}}=0.08$. Finally, the $\Delta_{47}$ of each sample was corrected based on the average offset from the accepted value of the $\Delta_{47}$ of two lab standards measured during the same week as the sample.

$\Delta_{47}$ values were converted to water temperature using the following equation \citep{Ghosh2006}:

\begin{align}
\Delta_{47}=50.0592\left(10^6\text{T}^{-2}\right)-0.02
\end{align}

\noindent where T is temperature in degrees Kelvin.

The \deltao O of water was calcuated from the temperature and the measured \deltao O of the sample using the temperature-dependent fractionation, $\alpha$, of oxygen isotopes between water and calcite \citep{Kim1997}:

\begin{align}
1000\ln\alpha=18.03\left(10^3\text{T}^{-1}\right)-32.42
\end{align}

\noindent where T is temperature in degrees Kelvin.

>>>>

Each sample was measured between one and three times. Approximately 10 milligrams of powdered sample were dissolved in phosphoric acid at 90\degrees C, which produces CO$_2$ with a known oxygen isotopic offset from the sample carbonate. The subsequent cleaning of the CO$_2$ followed procedures outlined by \cite{Ghosh2006}. The cleaned CO$_2$ was analyzed on a Finnigan MAT253 gas source mass spectrometer configured to collect masses 44 through 49. The sample gas was measured in eight separate acquisitions, each of which was bracketed by a measurement of reference CO$_2$ with known \deltac C and \deltao O. The \deltac C, \deltao O, $\delta_{45}$, $\delta_{46}$, $\delta_{47}$, and $\delta_{48}$ were calculated for each sample relative to the reference CO$_2$. Acquisition-to-acquisition standard deviations were on average 0.03$\permil$ for $\delta_{47}$, 0.04$\permil$ for \deltao O, and 0.02$\permil$ for \deltac C. Raw measured values for $\Delta_{47}$ and $\Delta_{48}$ of each sample, defined as $\Delta_{47-[\text{SGvsWG}]}$ and $\Delta_{48-[\text{SGvsWG}]}$, were also calculated relative to the reference CO$_2$ following equations 3 and 4 in \cite{Huntington2009}.

All $\Delta_{47}$ values are projected into the absolute reference frame \citep{Dennis2011}. When possible, raw $\Delta_{47-[\text{SGvsWG}]}$ values are converted to the absolute reference frame using an empirical transfer function (ETF) constructed from the heated and equilibrated  gases measured during the same week as the samples. Heated gases are aliquots of CO$_2$ that were sealed in quartz tubes, heated at 1000\degrees C for three hours, and then cooled rapidly to room temperature. Equilibrated gases are aliquots of CO$_2$ that were equilibrated with water at 25\degrees C for at least 24 hours. Heated gases were made from two separate reservoirs of CO$_2$ with different \deltao O compositions, and equilibrated gases were equilibrated with either of two separate waters, also differing in their \deltao O compositions. Heated and equilibrated gases were cleaned following the same procedure as the samples. The ETF was constructed and then used to convert raw $\Delta_{47-[\text{SGvsWG}]}$ values following \cite{Dennis2011}. Finally, the resulting $\Delta_{47-\text{RF}}$ was corrected for the acid extraction at 90\degrees C by adding 0.092$\permil$ \citep{Henkes2013}.

Some samples were measured during weeks when no equilibrated gases were measured, so an ETF could not be constructed for those weeks. These samples were projected into the absolute reference frame using a secondary transfer function (STF). Accepted values of two California Institute of Technology intralab standards were determined in the absolute reference frame from measurements of these standards during weeks when an ETF could be constructed from heated and equilibrated gas measurements. To create the STF, the accepted values for the two standards, as well as the accepted value of 0.0266$\permil$ for heated gases, were plotted against their $\Delta_{47}$ as determined in the \cite{Ghosh2006} reference frame. $\Delta_{47}$ values of the samples in the \cite{Ghosh2006} reference frame were corrected using this STF following equation C1 in \cite{Dennis2011}. A separate STF was constructed for each week of measurements during which no equilibrated gases were measured.

Although all samples were cleaned prior to introduction into the mass spectrometer, the possibility for contamination from hydrocarbons and chlorocarbons still exists. To screen for contamination, the $\delta_{48}$ and $\Delta_{48}$ of each sample were compared to a line of best fit through a plot of $\delta_{48}$ versus $\Delta_{48}$ of the heated gases, all in the \cite{Ghosh2006} reference frame. Samples whose $\Delta_{48}$ fell more than 1$\permil$ off the line were discarded due to likely contamination.

$\Delta_{47-\text{RF-AC}}$ values were converted to temperature using the \cite{Ghosh2006} calibration line, which is projected into the absolute reference frame in equation 9 in \cite{Dennis2011}. The \deltao O of water was calcuated from the temperature and the measured \deltao O of the sample using the temperature-dependent fractionation, $\alpha$, of oxygen isotopes between water and calcite, $1000\ln\alpha=18.03\left(10^3\text{T}^{-1}\right)-32.42$, where T is temperature in degrees Kelvin \citep{Kim1997}.
